---
title: "Calculate landscape parameters from Ridgefield geoTIFF DEM"
output: 
  bookdown::html_document2:
    fig_width: 6.4
    fig_height: 4.8
    number_sections: no
    self_contained: no
    df_print: default
    fig_caption: yes
    smart: no
---

```{r include=FALSE}
load(".RData")
```

[**Note**: not all the code is shown.]{style="background-color: #FFFF00;"}

First load packages and some other setup stuff

```{r message=FALSE, warning=FALSE, results='hold'}
library(sf)
library(raster) # slope parameters from raster::terrain() function
library(rgdal)
library(ggplot2)
library(ggpubr)
library(viridis)
library(fgeo.analyze)
GDA94 <- CRS("+proj=geocent +ellps=GRS80 +units=m +no_defs +type=crs")
```

Create matrix of test elevations in Excel using summed exponential functions & import into R data frame `testElev`, check with plot:

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
# testElev <- read.table("clipboard", sep="\t", header=TRUE) # no accidental run
image(seq(498000,499950,50), seq(6405700,6407650,50),
      as.matrix(testElev[,2:41]))
contour(seq(498000,499950,50), seq(6405700,6407650,50),
        as.matrix(testElev[,2:41]), add=TRUE, labcex=1.)
```

Make into a 3-column dataframe (x,y,elev)

```{r}
elevDF <- data.frame(x=rep(seq(498000,499950,50),40),
                     y=rep(seq(6405700,6407650,50),each=40),
                     elev=c(as.matrix(testElev[,2:41])))
summary(elevDF)
```

Make RasterLayer from this 3-column dataframe, and add a CRS
```{r}
testRast <- raster::rasterFromXYZ(elevDF)
raster::crs(testRast) <- "+proj=utm +zone=50 +south"
```

Then calculate slope using `raster::terrain()`:

```{r}
test_slope <- raster::terrain(testRast, opt = "slope", unit = "degrees")
```

This section based on code at <https://datacarpentry.org/r-raster-vector-geospatial/01-raster-structure/>

Read Ridgefield geoTIFF into R rasterLayer

```{r eval=-1}
GDALinfo("Ridgefield_DEM_gda94_utm50_epsg28350.tif") # just file info
RF_elev <- raster("Ridgefield_DEM_gda94_utm50_epsg28350.tif")
  # also make a raster subset with no NA
  NWrect <- extent(498000,500000,6405700,6407000)
  NW_elev <- raster::crop(RF_elev,NWrect)
  RF_elevDF <- as.data.frame(RF_elev, xy=TRUE)
  colnames(RF_elevDF)[3] <- "elev"
```

# plot the data frame made from elevation raster
```{r fig.height=9, fig.width=10.5, message=FALSE, warning=FALSE}
elevMapGG <- ggplot() +
  geom_raster(data = RF_elevDF , aes(x = x, y = y, fill = elev)) +
  geom_sf(data = rf_hydrology, col="steelblue2") + # add the streams
  scale_fill_viridis_c(na.value = "#FFFFFF00", name="Elevation (m)") +
  labs(x = "Longitude", y="Latitude") +
  coord_sf(crs = st_crs(28350)) +
  annotate(geom="text", x=495000, y=6407500,
           label=paste0("Elevations based on DEM with 13m cells",
                        "\nProjection: Long-Lat using GDA94"),
           size=6, hjust = 0) +
  theme_bw() +
  theme(axis.title = element_text(size=16, face="bold"),
        axis.text = element_text(size=14),
        legend.title = element_text(size=14, face="bold"),
        legend.text = element_text(size=14))
elevMapGG
```

For the `terrain()` function see <https://arc2r.github.io/book/Surface.html>

## Working with slope

Extract slope from elevation data

```{r}
RF_slope <- terrain(RF_elev, opt = "slope", unit = "degrees", neighbors=4)
RF_slope@crs@projargs ; bbox(RF_slope) ; summary(RF_slope)
```

Make a data frame version of slope

```{r}
RF_slopeDF <- as.data.frame(RF_slope_MGA, xy=TRUE)
summary(RF_slopeDF)
```

Plot the raw slope data

```{r fig.height=6, fig.width=7, message=FALSE, warning=FALSE}
ggplot() +
  geom_raster(data = RF_slopeDF , aes(x = x, y = y, fill = slope)) +
  scale_fill_viridis_c(na.value = "#00000000") +
  labs(x = "Easting", y="Northing") +
  coord_fixed() +
  theme_bw()
```

Categorise the slopes into range classes and summarise \\
(nice breaks between classes found by trial and error!)

```{r}
RF_slopeDF$SlopeCat <- cut(RF_slopeDF$slope,
                       breaks = c(0,1,2,3,4,6,10,15,20,25,99),
                       labels=c("0-1","1-2","2-3","3-4", "4-6","6-10","12-15",
                                "15-20","20-25",">25"))
# summarise the slope categories
cat("Percentage of pixels in each slope category:")
(pctslopes <-
  round(100*table(RF_slopeDF$SlopeCat)/sum(table(RF_slopeDF$SlopeCat)),2))
```

# plot the categorised slope map
```{r fig.height=9, fig.width=10.5, message=FALSE, warning=FALSE}
slopeCatMapGG <- ggplot() +
  geom_raster(data = RF_slopeDF , aes(x = x, y = y, fill = SlopeCat)) +
  geom_sf(data = rf_hydrology, col="steelblue2") + # add streams
  scale_fill_viridis_d(na.value = "#FFFFFF00", name="Slope range (\u00B0)") +
  labs(x = "Longitude", y="Latitude") +
  coord_sf(crs=st_crs(28350)) +
  annotate(geom="text", x=495000, y=6407500,
          label=paste0("Slope categories based on DEM with 13m cells",
                        "\nProjection: Long-Lat using GDA94"),
           size=6, hjust = 0) +
  theme_bw() +
  theme(axis.title = element_text(size=16, face="bold"),
        axis.text = element_text(size=14),
        legend.title = element_text(size=14, face="bold"),
        legend.text = element_text(size=14))
slopeCatMapGG
```

---------------------------------------------

## Working with aspect

Extract aspect from elevation data

```{r}
RF_aspect <- terrain(RF_elev, opt = "aspect", unit = "degrees")
```

Make data frame version of aspect data

```{r}
RF_aspectDF <- as.data.frame(RF_aspect, xy=TRUE)
```

Map aspect with ggplot

```{r fig.height=9, fig.width=10.5, message=FALSE, warning=FALSE}
RF_aspectGG <- ggplot() +
  geom_raster(data = RF_aspectDF, aes(x = x, y = y, fill = aspect)) +
  # make interpretable colour gradient
  scale_fill_gradientn(na.value = "#00000000",
                       colors = c("cornsilk","gold","royalblue2","blue4",
                                  "blue4","royalblue2","gold","cornsilk"),
                       name="Aspect\ndirection (\u00B0)") +
  geom_sf(data = rf_hydrology, col="coral", size=0.8) +
  labs(x = "Longitude", y="Latitude") +
  annotate(geom="text", x=495000, y=6407500,
           label=paste0("Slope aspect directions based on DEM with 13m cells",
                        "\n(yellow shades ~ North, blue shades ~ South)",
                        "\nProjection: Long-Lat using GDA94"),
           size=6, hjust = 0) +
  coord_sf(crs=st_crs(28350)) +
  theme_bw() +
  theme(axis.title = element_text(size=16, face="bold"),
        axis.text = element_text(size=14),
        legend.title = element_text(size=14, face="bold"),
        legend.text = element_text(size=14))
RF_aspectGG
```

------------------------------------------------------------

## Terrain water flow direction

Extract flow direction from elevation data

```{r}
RF_flowdir <- terrain(RF_elev, opt = "flowdir")
```

Make data frame version of flow directions

```{r}
RF_flowdirDF <- as.data.frame(RF_flowdir, xy=TRUE)
```

Flow direction map with ggplot

```{r fig.height=9, fig.width=10.5, message=FALSE, warning=FALSE}
RF_flowdirDF$flowQ <- cut(RF_flowdirDF$flowdir, breaks=c(0,2^seq(0,7)+0.5),
                          labels=c("E","SE","S","SW","W","NW","N","NE"))
RF_flowGG <- ggplot() +
  geom_raster(data = RF_flowdirDF, aes(x = x, y = y, fill = flowQ)) +
  scale_fill_viridis_d(na.value = "#00000000",
                       option = "rocket",
                       name="Flow\ndirection") +
  geom_sf(data = rf_hydrology, col="coral", size=0.8) +
  labs(x = "Longitude", y="Latitude") +
  annotate(geom="text", x=495000, y=6407500,
           label=paste0("Water flow direction classes based on DEM with 13m cells",
                        "\nProjection: Long-Lat using GDA94"),
           size=6, hjust = 0) +
  coord_sf(crs=st_crs(28350)) +
  theme_bw() +
  theme(axis.title = element_text(size=16, face="bold"),
        axis.text = element_text(size=14),
        legend.title = element_text(size=14, face="bold"),
        legend.text = element_text(size=14))
RF_flowGG
```

--------------------------------------------------

## Writing data back to geoTIFF

Write the `.tif` files

```{r eval=FALSE}
writeRaster(RF_slope, filename="Ridgefield_Slope_EPSG28350.tif", format="GTiff")
writeRaster(RF_aspect, filename="Ridgefield_Aspect_EPSG28350.tif", format="GTiff")
writeRaster(RF_flowdir, filename="Ridgefield_flowdir_EPSG28350.tif", format="GTiff")
```

Check files with `rgdal::GDALinfo()`

```{r}
GDALinfo("Ridgefield_Slope_EPSG28350.tif") # just file info
GDALinfo("Ridgefield_Aspect_EPSG28350.tif") # just file info
GDALinfo("Ridgefield_flowdir_EPSG28350.tif") # just file info
```

--------------------------------------------------

not sure if this works! \
trying on the basis that slope curvature ~ second derivative of elevation???

```
RF_curvature <- terrain(RF_slope, opt = "slope", unit="degrees")
plot(RF_curvature, col=cividis(128))
```

## Try `fgeo.analyze::` R package for curvature

Try the test data first (made near top of this file)
```{r fig.height=7, fig.width=6, eval=-1}
elevTopoTib <- fgeo_topography(elevDF,gridsize=4,xdim=40,ydim=40, edgecorrect = F)
elevTopo <- as.data.frame(elevTopoTib)
image(seq(1,40), seq(1,40), as.matrix(testElev[,2:41]),xlab="x",ylab="y")
contour(seq(0,36,4), seq(0,36,4), matrix(elevTopo$convex,ncol=10),add=T,
        labcex=2)
# -=-=-=- OR -=-=-=-
image(seq(0,36,4), seq(0,36,4), matrix(elevTopo$convex,ncol=10),
      xlim=c(0,40), ylim=c(0,40))
elevTopo$curv <- cut(elevTopo$convex, breaks=c(-999999,0,999999),
                     labels=c("convex","concave"))
ggplot() +
  geom_raster(data = elevTopo, aes(x = gx, y = gy, fill = curv)) +
  scale_fill_manual(values = c("steelblue","wheat"),na.value = "#FFFFFF00") +
  coord_equal() +
  theme_bw()

test_slopeDF <- as.data.frame(test_slope, xy=T)
```

--------------------------------------------------

```
fgeo_topography() function still not working on Ridgefield data :( ####
even when using a raster subset with no NA values
NW_elevDF <- as.data.frame(NW_elev, xy=TRUE)
names(NW_elevDF)[3] <- "elev"
summary(NW_elevDF)
```

> NW_topo_pars <- fgeo_topography(NW_elevDF, gridsize=20, xdim=425, ydim = 360)

`Error: cannot allocate vector of size 59.7 Gb`

> NW_topo_pars <- fgeo_topography(NW_elevDF, gridsize=40, xdim=425, ydim = 360)

`Error in if (theta1 <= theta2) { : missing value where TRUE/FALSE needed`

> NW_topo_pars <- fgeo_topography(NW_elevDF, gridsize=40, xdim=425, ydim = 360,
                                edgecorrect = F)
                                
`Error in if (theta1 <= theta2) { : missing value where TRUE/FALSE needed`
