---
title: "Ridgefield to 2018 A horizon spatial analyses"
output: 
  bookdown::html_document2:
    fig_width: 6.4
    fig_height: 4.8
    number_sections: no
    self_contained: no
    df_print: default
    fig_caption: yes
    smart: no
---

[**Note**: not all the code is shown.]{style="background-color: #FFFF00;"}

## use these packages / options

```{r load packages and environment but do not show, message=FALSE, warning=FALSE}
library(sp)
library(lctools)
library(gstat)
library(sf)
library(maptiles)
library(prettymapr)
library(fields)
library(viridis)
library(ncf)
library(flextable)
  set_flextable_defaults(font.family = 'sans', theme_fun = "theme_zebra", 
      font.size = 11, text.align = "right",padding.left = 1, padding.right = 1,
      padding.top = 2, padding.bottom = 2)
```

## Read the data

```{r message=FALSE, warning=FALSE}
git <- "https://raw.githubusercontent.com/Ratey-AtUWA/ridgefield-2022/main/"
```

```{r importing the data, message=FALSE, warning=FALSE, results='hold'}
rf2022 <- read.csv(paste0(git,"rf2022.csv"), stringsAsFactors = TRUE)
rf22a <- droplevels(subset(rf2022, subset=rf2022$Horiz_simp=="A"))
rf18a <- read.csv(paste0(git,"rf2018A.csv"), stringsAsFactors = TRUE)
rf_boundary <- read.csv(paste0(git,"rf_boundary.csv"))
farmtrack <- read.csv(paste0(git,"farmtrack.csv"))
minor <- read.csv(paste0(git,"minorStrm.csv"))
nimb <- read.csv(paste0(git,"NimbedillingBrk.csv"))
rf_annot <- list(farmtrack = farmtrack, minor = minor, 
                 nimb = nimb, rf_boundary = rf_boundary)
```

# Making our base map of Ridgefield 

```{r define CRS, warning=FALSE, error=FALSE, results='hold'}
LongLat <- CRS("+proj=longlat +ellps=WGS84 
           +datum=WGS84 +no_defs") # uses Earth ellipsis specs from WGS84 datum
UTM50S <- CRS("+proj=utm +zone=50 +south") # just for Zone 50, S hemisphere!
```

### ├ Map extent object

```{r make map extent object, warning=FALSE, error=FALSE, results='hold'}
extent <- st_as_sf(x = data.frame(x = c(116.94,117.01), 
                                  y = c(-32.515,-32.465)), 
                   coords = c("x", "y"), crs = LongLat)
st_coordinates(extent)
extent_UTM <- st_transform(extent, crs=UTM50S)
st_coordinates(extent_UTM)
```

### ├ Getting and plotting the map tile data

```{r basemap, fig.height=7.05, fig.width=8, fig.align='center', fig.cap="Map of Ridgefield Farm and adjacent area (UTM projection, Zone 50S, EPSG:32750) used subsequently as the base map for spatial analyses. Generated using the maptiles R package, with ESRI WorldTopoMap tiles.", out.width="75%", warning=FALSE, message=FALSE, results='hide'}
rftiles <- get_tiles(extent_UTM, provider = "Esri.WorldImagery", 
                     crop = TRUE, zoom = 14) # make map object

par(oma=c(3,3,1,1), mar=c(4,4,1.5,1.5), mgp=c(1.4,0.3,0), 
    lend=2, ljoin=1, tcl=0.3)
plot(rftiles)
axis(1)
mtext(side=1, line=2, text="Easting (UTM Zone 50, m)",
      font=2, cex=1.2)
axis(2)
mtext(side=2, line=2, text="Northing (UTM Zone 50, m)",
      font=2, cex=1.2)
source(paste0(git,"rf_map_annot.R"))
with(rf22a, points(Easting, Northing, pch = 21, cex = 0.9, 
                   bg = "cyan", col="white"))
with(rf18a, points(Easting, Northing, pch = 21, cex = 0.9, bg = "wheat"))
addnortharrow(pos="topright", border=1, lwd=1, text.col="white",  
              padin=c(0.1,0.2), scale=1.2)
addscalebar(plotepsg = 32750, linecol = "black", label.col = "white",
            widthhint = 0.3, htin = 0.15, label.cex = 1.3)
legend("left", bty="o", bg = "grey33", inset = 0.01, 
       legend = c("Previous data","New up to 2022"), col=c("black","white"),
       pch = 21, pt.bg = c("wheat","cyan"), text.col = "white")
box()
```

## Spatial Autocorrelation

### ├ Calculate Global Moran's I

```{r global Moran, echo=TRUE, message=FALSE, warning=FALSE, results='hold'}
var0 <- "C.pct" # choose the variable of interest

# Calculate global Moran's I
data_temp <- na.omit(rf22a[,c("Easting", "Northing", var0)])
Coords <- cbind(data_temp$Easting, data_temp$Northing)
bw <- 8 # bw = bandwidth, = number of nearest point for calculation
mI <- moransI(Coords, bw, data_temp[,3])
cat("Global Moran's I for ",var0," from Ridgefield A horizon\n", sep="");
t(as.matrix(mI[c(2,4:7)]))
rm(list = c("data_temp","Coords","bw","mI","var0")) # remove temporary objects
```

### ├ Plot local Moran's I

```{r local-Moran, message=FALSE, warning=FALSE, fig.height=7.15, fig.width=10, fig.align='center', out.width="80%", fig.cap="Map of Local Moran's I for TOC concentrations in Ridgefield A horizon soils to 2022. The Global Moran's I parameter is also shown beneath the legend.", results='hide'}
palette("default");palette(c(palette(),"gray92","white","transparent"))
var0 <- "C.pct" # choose the variable of interest
data_temp <- na.omit(rf22a[,c("Easting", "Northing", var0)])
Coords <- cbind(data_temp$Easting, data_temp$Northing)
mI <- moransI(Coords, 8, data_temp[,3]) # log10 minimises skewness
local_moran <- l.moransI(Coords, 8, data_temp[,3], scatter.plot = FALSE)
plotdata <- data.frame(Easting=Coords[,1], Northing=Coords[,2], 
                       MoranI=local_moran$Ii, p_value=local_moran$p.value)
pos0 <- subset(plotdata, plotdata$MoranI>0 & plotdata$p_value<=0.05)
neg0 <- subset(plotdata, plotdata$MoranI<0 & plotdata$p_value<=0.05)
#
par(oma=c(3,3,1,1), mar=c(4,4,1.5,1.5), mgp=c(1.4,0.2,0), 
    lend=2, ljoin=1, tcl=0.3, lwd = 1)
layout(matrix(c(1,1,1,1,2),nrow = 1))
plot(rftiles)
axis(1)
mtext(side=1, line=1.7, text="Easting (UTM Zone 50, m)",
      font=2, cex=1.2)
axis(2)
mtext(side=2, line=1.7, text="Northing (UTM Zone 50, m)",
      font=2, cex=1.2)
addnortharrow(pos="topright", border=1, lwd=1, text.col=10,  
              padin=c(0.1,0.2), scale=1.2)
addscalebar(plotepsg = 32750, linecol = "white", label.col = "white",
            widthhint = 0.3, htin = 0.15, label.cex = 1.3)
box()
points(Coords, pch=3, cex=0.5, col = "lemonchiffon")
with(pos0, symbols(Easting, Northing, circles = 250*sqrt(MoranI*0.02), lwd=2, 
                   inches = F, fg = "cyan", bg = "#00808080", add= TRUE))
with(neg0, symbols(Easting, Northing, squares = 250*sqrt(MoranI*-0.04), lwd=2, 
                   inches = F, fg = "plum2", bg = "#c0404080", add= TRUE))
plot(c(0,1),c(0,1),type="n",bty="n",axes=F)
legend("top", bty = "o", cex = 1.5, title.cex = 1.5,
       legend = c("Positive I", "Negative I",NA, NA,NA,NA),
       title = paste0(" Local Moran's I for \n",var0, 
                      " in rf22a data \n(p \u2264 0.05 only)"),
       pch = c(21, 22,NA,NA,NA), pt.cex = c(3, 2.5,NA,NA,NA,NA),
       col = c("cyan", "plum2"), pt.bg = c("#00808080", "#C0404080"),
       text.col = "white", title.col = "pink", y.intersp = 0.9, bg="#394739")
text(0.5, 0.815, col = 10, cex = 1.2, 
     labels = paste("Global Moran's I:\n", signif(as.numeric(mI[2]),3),
                           "\nRandomization p:\n", signif(as.numeric(mI[7]),3)))
```

### ├ Plot 'LISA'

```{r lisa-plot, message=FALSE, warning=FALSE, fig.height=7.15, fig.width=10, fig.align='center', out.width="80%", fig.cap="Map of LISA for C.pct concentrations in the Ridgefield A horizon soil data to 2022.", results='hide',echo=FALSE}
data_temp <- na.omit(as.data.frame(cbind(rf22a$Easting,rf22a$Northing,
                                         rf22a$C.pct)))
colnames(data_temp) <- c("Easting","Northing","C.pct")


Coords <- cbind(data_temp$Easting, data_temp$Northing)
bw <- 8
locMI <- l.moransI(Coords,bw,data_temp$C.pct, scatter.plot = F)
LISA <- as.data.frame(cbind(Coords[,1:2], log10(data_temp$C.pct), 
                            locMI$Ii, 
                            locMI$p.value))
colnames(LISA) <- c("Easting", "Northing", "C.pct", "MoranI", "p_value") # ;head(LISA)
medC.pct <- median(LISA$C.pct, na.rm = TRUE)
HiHi <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI>0&LISA$C.pct>=medC.pct)
LoLo <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI>0&LISA$C.pct<medC.pct)
HiLo <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI<0&LISA$C.pct>=medC.pct)
LoHi <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI<0&LISA$C.pct<medC.pct)
#
par(oma=c(3,3,1,1), mar=c(4,4,1.5,1.5), mgp=c(1.4,0.2,0), 
    lend=2, ljoin=1, tcl=0.3, lwd = 1)
layout(matrix(c(1,1,1,1,2),nrow = 1))
palette(c("black","#FF8080","#8080FF","orange","skyblue",
          "#B00000","#0000B0","#FFA50040","#87CEEB40","white"))
plot(rftiles)
axis(1)
mtext(side=1, line=1.6, text="Easting (UTM Zone 50, m)",
      font=2, cex=1.2)
axis(2)
mtext(side=2, line=1.6, text="Northing (UTM Zone 50, m)",
      font=2, cex=1.2)
addnortharrow(pos="topright", border=1, lwd=1, text.col=10,  
              padin=c(0.1,0.2), scale=1.2)
addscalebar(plotepsg = 32750, linecol = "white", label.col = "white",
            widthhint = 0.3, htin = 0.15, label.cex = 1.5)
box()
with(data_temp, points(Easting, Northing, pch=3, cex = 1, col="lemonchiffon"))
with(HiHi, points(Northing ~ Easting, pch = 22,
                  col = 2, bg = 6, lwd = 2, cex = 2.7))
with(LoLo, points(Northing ~ Easting, pch = 21, 
                  col = 3, bg = 7, lwd = 2, cex = 3))
with(HiLo, points(Northing ~ Easting, pch = 25,  
                  col = 2, bg = 7, lwd = 2, cex = 2.4))
with(LoHi, points(Northing ~ Easting, pch = 24,
                  col = 3, bg = 6, lwd = 2, cex = 2.4))
plot(c(0,1),c(0,1),type="n",bty="n",axes=F)
legend("top", bty = "o", inset = 0.015, cex = 1.5, ncol = 1,
       legend = c("High-High", "Low-Low","High-Low","Low-High","Sample points"),
       pch = c(22, 21, 25,24,3), pt.cex = c(2.7,3,2.4,2.4,1), 
       pt.lwd = c(2,2,2,2,1),
       col = c(2,3,2,3,"lemonchiffon"), pt.bg = c(6,7,7,6,NA), 
       box.col = "#b0b0b0", box.lwd = 1.3, bg = "#384838", 
       title="\nSpatial\nautocorrelation\nclasses\nfor C.pct",
       title.col = "lemonchiffon", text.col=10, y.intersp = 1.1)
```

### ├ LISA maps for multiple variables

```{r multi-LISA, fig.height=10, fig.width=7.8, fig.align='center', fig.cap="LISA autocorrelation maps for C.pct, Al, Fe, S, As, Cu, Pb, and Zn concentrations in the Ridgefield A horizon soil data to 2022.", message=FALSE, warning=FALSE, echo=FALSE}
par(mfrow=c(3,2), mar=c(0,0,0,0), oma=c(3,3,0.5,0.5), mgp=c(1.6,0.2,0), 
    tcl=0.3,lend="square", ljoin="mitre")
palette(c("black","#FF8080","#8080FF","orange","skyblue",
          "#B00000","#0000B0","#FFA50040","#87CEEB40","white"))
varz <- c("Gravel.pct","C.N.ratio","pH.CaCl2","EC","CEC","ex.Na")
for (i in 1:length(varz)){
  data_temp <- na.omit(as.data.frame(cbind(rf22a$Easting,rf22a$Northing,
                                         rf22a[,varz[i]])))
  colnames(data_temp) <- c("Easting","Northing",varz[i])

  Coords <- cbind(data_temp$Easting, data_temp$Northing)
  bw <- 8
  locMI <- l.moransI(Coords,bw,data_temp[,varz[i]], scatter.plot = F)
  LISA <- as.data.frame(cbind(Coords[,1:2], log10(data_temp[,varz[i]]), 
                            locMI$Ii, 
                            locMI$p.value))
  colnames(LISA) <- c("Easting", "Northing", var0, "MoranI", "p_value")
  medC.pct <- median(LISA[,var0], na.rm = TRUE)
  HiHi <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI>0&LISA[,var0]>=medC.pct)
  LoLo <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI>0&LISA[,var0]<medC.pct)
  HiLo <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI<0&LISA[,var0]>=medC.pct)
  LoHi <- subset(LISA, LISA$p_value<=0.05&LISA$MoranI<0&LISA[,var0]<medC.pct)

  plot(rftiles)
  axis(1, labels=F,col=10); axis(2, labels = F, col=10)
  if(i > 4.5) {
    axis(1)
    mtext(side=1, line=1.6, text="Easting (UTM Zone 50, m)", font=2, cex=1)
    }
  if(round(i/2,0) != (i/2)) {
    axis(2)
    mtext(side=2, line=1.6, text="Northing (UTM Zone 50, m)", font=2, cex=1)
  }
  addnortharrow(pos="topleft", border=1, lwd=1, text.col=10,  
              padin=c(0.05,0.05), scale = 0.8)
  addscalebar(plotepsg = 32750, label.cex = 1.2, label.col = 10)
  box(col=10)
  with(data_temp, points(Easting, Northing, pch=3, cex=0.7, col="lemonchiffon"))
  with(HiHi, points(Northing ~ Easting, pch = 22,
                  col = 2, bg = 6, lwd = 2, cex = 1.8))
  with(LoLo, points(Northing ~ Easting, pch = 21, 
                  col = 3, bg = 7, lwd = 2, cex = 2))
  with(HiLo, points(Northing ~ Easting, pch = 25,  
                  col = 2, bg = 7, lwd = 2, cex = 1.8))
  with(LoHi, points(Northing ~ Easting, pch = 24,
                  col = 3, bg = 6, lwd = 2, cex = 1.8))
  mtext(varz[i],3,-1.55,adj=0.985,font=2)
  mtext(varz[i],3,-1.5,adj=0.98,font=2,col="white")
  legend(494200, 6407030, bty = "n", cex = 1.2, ncol=1, inset = 0.01,
       legend = c("Hi-Hi", "Lo-Lo","Hi-Lo","Lo-Hi","Samples"),
       pch = c(22, 21, 25,24,3), pt.cex = c(1.8,2,1.8,1.8,0.7), 
       pt.lwd = c(2,2,2,2,1),
       col = c(2,3,2,3,"lemonchiffon"), pt.bg = c(6,7,7,6,NA), text.col = 10,
       box.col = NA, box.lwd = 1.3, bg = "#00000080", 
       title=expression(italic("Key")), x.intersp = 0.8, 
       title.col = "white")

}
par(mfrow=c(1,1), mar=c(3,3,0.5,0.5), oma=c(0,0,0,0), mgp=c(1.6,0.2,0), 
    tcl=0.3,lend="square", ljoin="mitre")

```

```{r reset-layout, echo=FALSE}
layout(1)
```

```{r remove temps 2, message=FALSE, warning=FALSE, include=FALSE}
rm(list = c("data_temp","Coords","mI","locMI", "plotdata", "pos0", "neg0",
            "LISA","medC.pct","HiHi","LoLo","HiLo","LoHi","bw","i","local_moran",
            "var0","varz"))
```

<hr>

## Using the gstat package for geostatistics: variograms, kriging, and visualization

### ├ Make `sf` & `SpatialPointsDataFrame` objects from a data frame

In this example (as for Moran's I) we $log_{10}$-transform our variable if its
distribution is highly positively skewed. (Using the untransformed variable
would result in too many apparent upper outliers.)

```{r make SpatialPointsDataFrame object, paged.print=FALSE, results='hold'}
data0 <- na.omit(rf22a[,c("Easting","Northing","C.pct")])
data0[,3] <- log10(data0[,3])
C.pct_sf <- st_as_sf(data0, coords=c("Easting","Northing"), crs = st_crs(32750))
C.pct_sp <- as_Spatial(C.pct_sf)
summary(C.pct_sp)
```

### ├ Plot the spatial object for checking

To quickly check our data, we use the function `bubble()` from the
`sp` package to make a *bubble map* of our variable, where the symbol *area* is
proportional to the variable value (in this case, soil TOC concentration).

```{r spplot-spdf, fig.height=5, fig.width=8, fig.align='center', out.width="80%", fig.cap="Visualization of spatial point data object for log10-transformed C.pct concentrations in Ridgefield A horizon soil to 2022.", message=FALSE, warning=FALSE, results='hold'}
bubble(C.pct_sp, col=c("#ff000088", "#0000ff88"), main = "", 
       scales = list(draw = TRUE), xlab = "Easting", ylab = "Northing")
```

### ├ Plot a map with range-class symbols

```{r TOC-tukey-plot, echo=2:19, fig.width=8, fig.height=7.05, fig.align='center', fig.cap="Map of C.pct concentrations expressed as symbols for concentration ranges (UQR is 75th percentile, upWsk is upper whisker, med is median, LQR is 25th percentile, loWsk is lower whisker). Data are from Ridgefield A horizon soils to 2022.", out.width="80%", message=FALSE, warning=FALSE,echo=FALSE}
palette(c("black",rev(c("#EE0000C0", "#EE7621C0", "#EEC900C0", "#40E0D0C0", 
                        "#4870F0C0", "#9932CCC0")),"grey60","grey92", "white"))
par(oma=c(4,4,1.5,1.5), mgp=c(1.4,0.3,0), lend=2, ljoin=1, tcl=0.3)
plot(rftiles)
axis(1)
mtext(side=1, line=2, text="Easting (UTM Zone 50, m)",
      font=2, cex=1.2)
axis(2, mgp=c(2, 0.5, 0))
mtext(side=2, line=2, text="Northing (UTM Zone 50, m)",
      font=2, cex=1.2)
addnortharrow(pos="topright", border=1, lwd=1, text.col=1,  
              padin=c(0.2,0.2), scale=0.8)
addscalebar(plotepsg = 32750, htin = 0.16, label.cex = 1.2, pos = "bottomright")
# construct and plot range classes in Tukey boxplot thresholds
C.pct_sp@data$tukey <- cut(C.pct_sp@data$C.pct, 
    breaks = c(0,boxplot.stats(C.pct_sp@data[,1])$stats,999999),
    labels = c("-outliers","loWx-LQR","LQR-med","med-UQR","UQR-upWx","+outliers"))
polygon(rf_boundary[,1:2], lty=3, lwd=2, border="khaki")
points(C.pct_sp@coords, pch = 21, col = 10,
       cex=(c(.5,.65,.85,1,1.4,1.8)*2)[C.pct_sp@data$tukey], 
       bg=seq(2,7)[C.pct_sp@data$tukey])
     
legend(494360, 6407050, legend = rev(c("-outliers","loWsk to LQR","LQR to med",
                                "med to UQR","UQR to upWsk","+outliers")), 
       cex = 1., pch=21, pt.cex=rev(c(.5,.65,.85,1,1.4,1.8))*2, inset = 0.07, 
       bty="n", y.intersp=1.15, pt.bg=seq(7,2,-1), x.intersp = .9,
       text.col = 10, col = 10)
text(399870, 6468410, 
     labels=expression(bold("C.pct ranges")), pos = 4, cex = 1)
box()
```

<p style="text-align: center;"><br>**Table 1: Bins for boxplot-categorised-point map in Figure \@ref(fig:TOC-tukey-plot):**</p>

```{r tukey-map-bins, echo=FALSE}
tukbins <- data.frame(bins=c("-outliers","loWsk to LQR","LQR to med",
                                "med to UQR","UQR to upWsk","+outliers"),
      logfrom=signif(c(-4,boxplot.stats(C.pct_sp@data[,1])$stats),3),
      logto=signif(c(boxplot.stats(C.pct_sp@data[,1])$stats,Inf),3),
      from=round(10^c(-4,boxplot.stats(C.pct_sp@data[,1])$stats),3),
      to=round(10^c(boxplot.stats(C.pct_sp@data[,1])$stats,Inf),3))
flextable(tukbins) |> autofit()
```

## Variograms and Kriging

### ├ Make a binned simple variogram object

```{r make-variog-obj, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Plot of experimental binned variogram for C.pct in Ridgefield A horizon soils to 2022.", fig.height=3.5, fig.width=4.5, out.width="50%"}
data0 <- na.omit(rf22a[,c("Easting","Northing","Gravel.pct")])
keeprows <- which(data0$Gravel.pct > 1e-6) # find nonzeros
data0 <- data0[keeprows,] # remove zeros
data0[,3] <- log10(data0[,3])
Gravel.pct_sf <- st_as_sf(data0, coords=c("Easting","Northing"), crs = st_crs(32750))
Gravel.pct_sp <- as_Spatial(Gravel.pct_sf)

par(mar = c(3,3,1,1), mgp = c(1.7,0.3,0), font.lab=2, tcl=0.3)

variog_Gravel.pct <- variogram(Gravel.pct~1, Gravel.pct_sf, cressie = TRUE) # 
plot(variog_Gravel.pct$dist, variog_Gravel.pct$gamma, xlab = "Distance (m)",
     pch=19, col="blue2", 
     ylab = expression(bold(paste("Semivariance (",gamma,")"))),
     ylim = c(0,max(variog_Gravel.pct$gamma)))
mtext(paste(attributes(variog_Gravel.pct)$what,"with",
            length(variog_Gravel.pct$dist),"bins"), side=1, line=-1.5, adj=0.98)
```

<p style="text-align: center;">**<br>Table 2: Data for Figure \@ref(fig:make-variog-obj)**.</p>

```{r print-variog-obj, paged.print=FALSE, echo=FALSE}
vgr <- variog_Gravel.pct
colnames(vgr)[1:3] <- c("n_points_in_bin","dist_metres","semivariance")
vgr[,2] <- round(vgr[,2],0)
vgr[,3] <- round(vgr[,3],3)
flextable(vgr[,1:3]) |> autofit(); rm(vgr)
```

### ├ Fit a variogram model using weighted least squares

```{r fit variogram model, fig.height=3, fig.width=4, message=FALSE, warning=FALSE, paged.print=FALSE, results='hold'}
vmodel_Gravel.pct <- fit.variogram(variog_Gravel.pct, 
                              model = vgm(psill = NA, model = "Exp", 
                                          nugget = 0.2, 
                                          cutoff = max(variog_Gravel.pct$dist)))
vmodel_Gravel.pct
```

```{r plot-exp-model-variograms, fig.height=3.6, fig.width=5, fig.align='center', fig.cap="Plot of experimental binned variogram, and exponential variogram model, for Gravel.pct in Ridgefield A horizon soils to 2022.", out.width="50%", results='hide', echo=FALSE}
par(mar = c(4,4,1,1), mgp = c(1.7,0.3,0), tcl=0.25, font.lab=2)
with(variog_Gravel.pct, plot(dist, gamma, ylim = c(0, max(gamma)), cex.lab=1.2, 
                   lwd=2, xlab = "Distance (m)", 
                   ylab = expression(bold(paste("Semivariance (",gamma,")")))))
# plot model
x0 <- variog_Gravel.pct$dist
y0 <- vmodel_Gravel.pct$psill[1] + 
  vmodel_Gravel.pct$psill[2]*(1 - exp(-1*x0/vmodel_Gravel.pct$range[2]))
lines(x0,y0,col=2)
legend("bottomright", bty = "n", cex=1, pch=c(1,NA), pt.lwd=2, lwd=c(NA,1), 
       col=c(1,2), pt.cex=1.1,legend = c("Empirical variogram",
                  paste(vmodel_Gravel.pct$model[2],"variogram model")))
text(max(x0),max(y0)/2,pos=2,offset=0, cex=0.75, family='mono', col = 2,
     labels=paste(
       " nugget =",signif(vmodel_Gravel.pct$psill[1],3),"\n",
       "partial sill =", signif(vmodel_Gravel.pct$psill[2],3),"\n",
       "range =", signif(vmodel_Gravel.pct$range[2],3),"m\n",
       "practical range =",signif(vmodel_Gravel.pct$range[2]*3,3),"m" 
  ))
# NOTE factor (range -> practical range) depends on model
```

### ├ Perform kriging interpolation

### ├ first make a grid mask

```{r make grid masked to AFR sampling zones, message=FALSE, warning=FALSE, results='hold'}
grid0 <- expand.grid(seq(round(min(Gravel.pct_sp@coords[,1]),1), 
                            round(max(Gravel.pct_sp@coords[,1]),1), 10),
                        seq(round(min(Gravel.pct_sp@coords[,2]),1), 
                            round(max(Gravel.pct_sp@coords[,2]),1), 10))
rfgrid <- SpatialPoints(grid0, proj4string = UTM50S)

irregpoly <- Polygon(rf_boundary[,c(1,2)], hole=F)
irregPolys = Polygons(list(irregpoly),1)
gridMask = SpatialPolygons(list(irregPolys), 
                           proj4string = UTM50S)
inOrOut <- as.vector(over(rfgrid, gridMask))
rfgrid <- rfgrid[which(inOrOut>0)]
cat("Prediction grid:\n"); summary(rfgrid)
```

### ├ Krige to grid
```{r krige-to-grid, results='hold'}
kriged_Gravel.pct <- krige(formula = Gravel.pct~1, locations = Gravel.pct_sp, 
                      newdata = rfgrid, model = vmodel_Gravel.pct)
summary(kriged_Gravel.pct)
cat("_____________________________________________________________________\n\n")
idw_Gravel.pct <- idw(formula = Gravel.pct~1, locations = Gravel.pct_sp, 
                      newdata = rfgrid, idp = 2)
idw_Gravel.pct@data$var1.var <- NULL
summary(idw_Gravel.pct)
```

### ├ Simple plot of kriging output

We can then us the spplot() function from sp to visualise the 
kriging predictions and variance, but without a background map.

```{r spplot-kriging-output, fig.width=8, fig.height=5.85, fig.align='center', echo=2, fig.cap="Plots of simple kriging predictions and variance for log-transformed Gravel.pct in Ridgefield A horizon soils to 2022.", out.width="75%", warning=FALSE, error=FALSE, results='hold'}
par(oma=c(0,0,0,0))
spplot(kriged_Gravel.pct, colorkey=T, pch=15, # with kriging variances
    names.attr = c("log10[Gravel.pct] kriging predictions",
                   "log10[Gravel.pct] kriging variance"), 
    scales = list(draw = TRUE), col.regions = viridis::mako(100),
    xlab="Longitude", ylab="Latitude")
```

### ├ Plot a map with overlay of the kriging predictions & kriging variance

```{r map-krig-mod-pred, fig.height=12.8, fig.width=7.8, fig.align='center', fig.cap="Map showing (a) predictions and (b) variance for kriging model of Gravel.pct in Ridgefield A horizon soils to 2022. Sample points are tiny dots.", message=FALSE, warning=FALSE}
par(mfrow=c(2,1), oma=c(4,4,1.5,1.5), mgp=c(1.4,0.3,0), lend=2, ljoin=1, tcl=0.3)
plot(rftiles)
axis(1, mgp=c(2, 0.3, 0), labels=F)
# mtext(side=1, line=1.5, text="Easting (UTM Zone 50, m)",
#       font=2, cex=1.2)
axis(2, mgp=c(2, 0.3, 0), cex.axis = 1)
mtext(side=2, line=1.5, text="Northing (UTM Zone 50, m)",
      font=2, cex=1.2)
axtx <- c(1,2,3,5,10,20,30,50,100,200,300,500,1000)
colgrad <- gsub("FF","98", viridis::mako(128)[round(seq(1, 128, l=6),0)])
rect(494500, 6402250, 501000, 6403000, col="#FFFFFF80", border=NA)
quilt.plot(rfgrid@coords[,1], rfgrid@coords[,2],kriged_Gravel.pct@data[,1],
           add = T, horizontal=T, 
           col = colorRampPalette(colgrad, alpha = TRUE)(64),
           legend.lab = expression(bold("Kriging prediction of Gravel.pct (mg/kg)")), 
           legend.mar = 6, legend.cex = 1, legend.line = 1.2, text.col=10,
           axis.args = list(at=log10(axtx), labels=axtx, mgp=c(1.2,0.1,0)))
addnortharrow(border=1, lwd=1, text.col=1, padin=c(0.1,0.1), scale = 0.9)
addscalebar(plotepsg = 32750, pos = "topleft")
points(rf22a[,c("Easting","Northing")], pch=16, cex=0.1, col = "gold")
box() ; mtext("(a)",3,-1.5, cex=1.5,col=10)

plot(rftiles)
axis(1, mgp=c(2, 0.3, 0), cex.axis = 1)
mtext(side=1, line=1.5, text="Easting (UTM Zone 50, m)",
      font=2, cex=1.2)
axis(2, mgp=c(2, 0.3, 0), cex.axis = 1)
mtext(side=2, line=1.5, text="Northing (UTM Zone 50, m)",
      font=2, cex=1.2)
axtx <- pretty(kriged_Gravel.pct@data[,2])
colgrad <- gsub("FF","80", viridis::turbo(128)[round(seq(1, 128, l=6),0)])
rect(494500, 6402250, 501000, 6403000, col="#FFFFFF80", border=NA)
quilt.plot(rfgrid@coords[,1], rfgrid@coords[,2],kriged_Gravel.pct@data[,2],
           add = T, horizontal=T, 
           col = colorRampPalette(colgrad, alpha = TRUE)(64),
           legend.lab = expression(bold("Kriging variance in Gravel.pct (mg/kg)")), 
           legend.mar = 6, legend.cex = 1, legend.line = 1.2, legend.col="white",
           axis.args = list(at=axtx, labels=axtx, mgp=c(1.2,0.1,0)))
addnortharrow(border=1, lwd=1, text.col=1, padin=c(0.1,0.1), scale = 0.9)
addscalebar(plotepsg = 32750, pos = "topleft")
points(rf22a[,c("Easting","Northing")], pch=16, cex=0.1, col="orchid")
box() ; mtext("(b)",3,-1.5, cex=1.5,col=10)
```

