---
title: "Ridgefield to 2018 A horizon spatial analyses"
knit: litedown:::knit
output: 
  html_document:
    fig_caption: yes
    self_contained: no
    number_sections: no
    toc: TRUE
    theme: cyborg
    highlight: kate
---

```{r author-date, echo=FALSE, results='asis'}
cat('<p id="top"><em>Andrew Rate, School of Agriculture and Environment, The University of Western Australia, ', as.character(Sys.Date()),'</em></p>')
```

<span style="background-color: #FFFF00;">**Note**: not all the code is shown.</span>

## use these packages / options

```{r load packages and options, message=FALSE, warning=FALSE}
library(gstat)
library(spdep)
library(sf)
library(maptiles)
library(prettymapr)
library(TeachingDemos)
library(ggplot2)
library(ggspatial)
library(shadowtext)
library(fields)
library(viridis)
library(ncf)
library(flextable)
  set_flextable_defaults(font.family = 'sans', theme_fun = "theme_zebra", 
      font.size = 11, text.align = "right",padding.left = 1, padding.right = 1,
      padding.top = 2, padding.bottom = 2)
```

## Read the data

```{r define-github-url, message=FALSE, warning=FALSE}
git <- "https://raw.githubusercontent.com/Ratey-AtUWA/ridgefield-2022/main/"
```

```{r importing the data, message=FALSE, warning=FALSE, results='hold'}
rf2022 <- read.csv(paste0(git,"rf2022.csv"), stringsAsFactors = TRUE)
rf22a <- droplevels(subset(rf2022, subset=rf2022$Horiz_simp=="A"))
rf18a <- read.csv(paste0(git,"rf2018A.csv"), stringsAsFactors = TRUE)
rf_boundary <- read.csv(paste0(git,"rf_boundary.csv"))
farmtrack <- read.csv(paste0(git,"farmtrack.csv"))
minor <- read.csv(paste0(git,"minorStrm.csv"))
nimb <- read.csv(paste0(git,"NimbedillingBrk.csv"))
rf_annot <- list(farmtrack = farmtrack, minor = minor, 
                 nimb = nimb, rf_boundary = rf_boundary)
rf_boundary <- rbind(rf_boundary,rf_boundary[1,])
rfb <- st_polygon(list(as.matrix(rf_boundary[,1:2]))) |> st_sfc(crs=32750)
localityRF <- st_read("locality-RF.kml") |> st_transform(crs=32750)
secret <- read.csv("../StadiaMaps-API.csv")[1] |> as.character()
# secret <- readLines("../thunderforest-API-key.txt")
```

# Making our base map of Ridgefield 

```{r define CRS, warning=FALSE, error=FALSE, results='hold'}
LongLat <- st_crs(4326) # uses Earth ellipsis specs from WGS84 datum
UTM50S <- st_crs(32750) # just for Zone 50, S hemisphere!
```

### ├ Map extent object and raster

```{r make map extent object, warning=FALSE, error=FALSE, results='hold'}
extent <- st_as_sf(x = data.frame(x = c(116.94,117.01), 
                                  y = c(-32.515,-32.465)), 
                   coords = c("x", "y"), crs = LongLat)
st_coordinates(extent)
extent_UTM <- st_transform(extent, crs=UTM50S)
st_coordinates(extent_UTM)
rfRast <- get_tiles(extent_UTM, provider = "Esri.WorldImagery", 
                     crop = TRUE, zoom = 14) # make map object
rfLocRast <- get_tiles(localityRF, provider="Stadia.StamenTerrain", crop=T,
                       apikey = secret, zoom=13, forceDownload = TRUE) #Stadia.StamenTerrain
rf_paddocks <-
  read_sf(paste0("../ridgefield-2022/","/Paddocks/Paddocks_gda94_utm50_EPSG28350.shp")) |> 
  st_transform(crs=st_crs(32750))
```

```{r make-boundary-gap, include=FALSE}
(EN0 <- unname(c(st_coordinates(extent_UTM)[1,],st_coordinates(extent_UTM)[2,])))
extLg <- matrix(c(floor(EN0[1]), floor(EN0[2]), floor(EN0[1]), ceiling(EN0[4]), 
                  ceiling(EN0[3]), ceiling(EN0[4]), ceiling(EN0[3]), floor(EN0[2]),
                  floor(EN0[1]), floor(EN0[2])), ncol=2, byrow = T)
rfGap <- st_polygon(list(extLg, as.matrix(rf_boundary[,1:2]))) |> st_sfc(crs=32750)
```

### ├ Plotting the map tile data

```{r locality-map-gg, fig.height=8.3, fig.width=12.4}
ggplot() +
  geom_sf(data=localityRF) +
  annotation_spatial(data=rfLocRast) +
  geom_sf(aes(fill=Name), data=rf_paddocks,color="#777") +
  scale_fill_viridis_d(begin=0.4, end=0.8, alpha=0.5, guide="none") +
  annotation_north_arrow(location="tr", style=north_arrow_fancy_orienteering) +
  annotation_scale(width_hint=0.22) +
  geom_shadowtext(aes(x=498090, y=6405073, label="Ridgefield\nFarm"), size=5,
                  color="#040", bg.color="#fff")+
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x="Longitude", y="Latitude", caption="CRS: WGS84 (EPSG:4326)") +
  theme_bw() +
  theme(axis.title = element_text(size=14, face=2),
        plot.caption = element_text(colour = "#777", face=3))
```

```{r make-sf-objs-of-A-horizon-data}
# make an sf data frame of A horizon data
rf22A.utm <- st_as_sf(rf22a[which(!is.na(rf22a$Northing)),], 
                    coords=c("Easting", "Northing"), crs=32750, remove=F)
rf18A.utm <- st_as_sf(rf18a, coords=c("Easting", "Northing"), crs=32750, remove=F)
```

```{r basemap, fig.height=7.4, fig.width=9, fig.align='center', fig.cap="Map of Ridgefield Farm and adjacent area (UTM projection, Zone 50S, EPSG:32750) used subsequently as the base map for spatial analyses. Generated using the maptiles R package, with ESRI WorldTopoMap tiles.", out.width="75%", warning=FALSE, message=FALSE}
rf1822 <- rbind(data.frame(Year=rep(2018,nrow(rf18A.utm)),
                           rf18A.utm[,c("Profile","Easting","Northing")]),
                data.frame(Year=rep(2022,nrow(rf22A.utm)),
                           rf22A.utm[,c("Profile","Easting","Northing")]))
rf1822$Year <- as.factor(rf1822$Year)
rf1822 <- st_as_sf(rf1822, coords=c("Easting", "Northing"), crs=32750)

ggplot()+
  geom_sf(data=extent_UTM, col="transparent") +
  annotation_spatial(data=rfRast) +
  labs(x="Easting (m)", y="Northing (m)", 
       caption="CRS: UTM Zone 50 South (EPSG:32750)") +
  geom_sf(data=rfGap, fill="#ffffff80", col="transparent") +
  geom_sf(aes(fill=Year), data=rf1822, shape=21, col="#FFF", size=3) +
  geom_sf(data=rf18A.utm, shape=21, fill="chocolate", col="#FFF", size=3.5, guide=F) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_manual(values=c("chocolate","wheat")) +
  annotation_north_arrow(location="tr", style=north_arrow_fancy_orienteering) +
  annotation_scale(location="bl", text_cex=1.1) +
  coord_sf(datum=st_crs(32750)) +
  theme_bw() + theme(axis.title=element_text(size=14, face="bold"),
                     legend.key=element_rect(fill="#8c8c85"))
```

```{r read-landforms-shapefiles, message=FALSE, warning=FALSE, results='hide'}
RF_Res_sf <- st_read("Ridgefield_residual.shp")
RF_Ero_sf <- st_read("Ridgefield_erosional.shp")
RF_Dep_sf <- st_read("Ridgefield_depositional.shp")
```

**Preliminary step 1**: make a depth-to-clay layer. The clay content used to define
"clay" is taken from the texture triangle used in the Australian Soil 
Classification (see the `soiltexture` R package):
- texture `%in%` {clay, silty clay, sandy clay}: %clay ≥ 26.5%
- texture `%in%` {clay loam, silty clay loam, sandy clay loam}: %clay ≥ 17%

```{r prelim-make-depthToClay, message=FALSE, warning=FALSE, paged.print=FALSE}
# make a list with same number of items as profiles with shallowest sample ≥ 17% clay
gt17 <- tapply(rf2022$Clay.pct, rf2022$Profile, FUN = function(x){min(which(x >= 17))})
gt17[gt17==Inf] <- NA # we get Inf values since some profiles never ≥ 17%
# find row numbers of first sample in profiles
profileBeg <- tapply(as.integer(row.names(rf2022)), rf2022$Profile, identity)
startRows <- unname(unlist(lapply(profileBeg,min)))

# use first sample in profile & shallowest sample ≥ 17% clay to make 
# depth to clay vector for each profile (some values will be NA)
d2c <- rf2022$Depth.upper[startRows+(gt17-1)]
# assign nominally large depth 
# if no texture for profile finer than sandyClayLoam (i.e. NA values)
# d2c[is.na(d2c)] <- 300
# or sample a distribution??? if too many values the same is problematic ??? 
d2c[is.na(d2c)] <- round(sample(rnorm(1e6,500,100),length(d2c[is.na(d2c)])))

# aggregate coordinates
profileE <- with(rf2022, aggregate(Easting, by=list(Profile), mean, na.rm=T))
profileN <- with(rf2022, aggregate(Northing, by=list(Profile), mean, na.rm=T))

profileCoords <- data.frame(Profile=profileE[,1],
                            Easting=profileE[,2],
                            Northing=profileN[,2],
                            DepthToClay=d2c)
(DTC <- st_as_sf(profileCoords, coords=c("Easting","Northing"),
                          crs=32750, remove=F))
```

## Spatial Autocorrelation

### ├ Calculate Global Moran's I

```{r global Moran, echo=TRUE, message=FALSE, warning=FALSE, results='hold'}
var0 <- "DepthToClay"
MI.data <- na.omit(DTC[,var0])
# then since `C.pct` has positively skewed distribution: (OMIT IF APPROPRIATE)
MI.data[,var0] <- st_drop_geometry(MI.data[,var0])

# make neighbour-weights object for 4 nearest neighbours
rfA.neighbWts <- knearneigh(MI.data, 4) |> knn2nb() |> nb2listw()
# Calculate global Moran's I
(rfA.MI <- moran.test(unlist(st_drop_geometry(MI.data[,var0])), rfA.neighbWts))

cat(paste0("🟨 Global Moran's I for ",var0," from Ridgefield soil profiles = ",
           signif(rfA.MI$estimate[1],3),", p = ",signif(rfA.MI$p.v[1],3),"\n"))
```

### ├ Plot local Moran's I

```{r local-Moran, message=FALSE, warning=FALSE, fig.height=7.4, fig.width=9, fig.align='center', out.width="80%", fig.cap="Map of Local Moran's I for TOC concentrations in Ridgefield A horizon soils to 2022. Only points where p-value for Local Morans I ≤ 0.05 are shown: red for positive, blue for negative. The Global Moran's I parameter is also shown beneath the legend.", results='hide'}

# calculate local Moran's I at each coordinate
rfA.LMI <- localmoran(unlist(st_drop_geometry(MI.data[,var0])), rfA.neighbWts)
MI.data$LMI <- as.data.frame(rfA.LMI)[,1]
MI.data$pLMI <- as.data.frame(rfA.LMI)[,5]
MI.data$LISA <- attributes(rfA.LMI)$quadr[,2]
MI.pos <- MI.data[MI.data$LMI>0 & MI.data$pLMI<=0.05,]
MI.neg <- MI.data[MI.data$LMI<=0 & MI.data$pLMI<=0.05,]

ggplot()+
  geom_sf(data=extent_UTM) +
  annotation_spatial(data=rfRast) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x="Easting (m)", y="Northing (m)", 
       caption=paste("Global Moran's I =",signif(rfA.MI$estimate[1],3),"\n",
                     "CRS: UTM Zone 50 South (EPSG:32750)")) +
  geom_sf(data=rfGap, fill="#ffffff80", col="transparent") +
  geom_sf(aes(size=LMI), data=MI.pos, shape=21, fill="#ff606080", col="#FFF") +
  geom_sf(aes(size=-1*LMI), data=MI.neg, shape=21, fill="#6060ff80", col="#FFF") +
  annotation_north_arrow(location="tr", style=north_arrow_fancy_orienteering) +
  annotation_scale(location="bl", text_cex=1.1) +
  theme_bw() + theme(axis.title=element_text(size=14, face="bold"),
        plot.caption = element_text(colour = "#777", face=3)) +
  coord_sf(datum=st_crs(32750))
```

### ├ Plot 'LISA'

```{r lisa-plot, message=FALSE, warning=FALSE, fig.height=7.4, fig.width=9, fig.align='center', out.width="80%", fig.cap="Map of median-based-LISA for C.pct concentrations in the Ridgefield A horizon soil data to 2022. Only points where p-value for Local Morans I ≤ 0.05 are shown.", results='hide',echo=FALSE}
MI.LISA <- MI.data[MI.data$pLMI<=0.05,]

ggplot()+
  geom_sf(data=extent_UTM) +
  annotation_spatial(data=rfRast) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x="Easting (m)", y="Northing (m)", 
       caption=paste("Global Moran's I =",signif(rfA.MI$estimate[1],3),"\n",
                     "CRS: UTM Zone 50 South (EPSG:32750)")) +
  geom_sf(data=rfGap, fill="#ffffff80", col="transparent") +
  annotation_north_arrow(location="tr", style=north_arrow_fancy_orienteering) +
  annotation_scale(location="bl", text_cex=1.1) +
  geom_sf(aes(shape=LISA, fill=LISA), data=MI.LISA, col="#FFF", size=3) +
  scale_shape_manual(values = c(25,23,23,24)) +
  scale_fill_viridis_d(option="plasma") +
  coord_sf(datum=st_crs(32750)) +
  theme_bw() +
  theme(legend.key = element_rect(fill="#8c8c85"))
```

<hr>

## Using the gstat package for geostatistics: variograms, kriging, and visualization

### ├ Make `sf` object from a data frame

In this example (as for Moran's I) we $log_{10}$-transform our variable if its
distribution is highly positively skewed. (Using the untransformed variable
would result in too many apparent upper outliers.)

```{r make-sf-point-object, paged.print=FALSE, results='hold'}
data0 <- na.omit(rf22a[,c("Easting","Northing","C.pct")])
data0[,3] <- log10(data0[,3])
C.pct_sf <- st_as_sf(data0, coords=c("Easting","Northing"), crs = st_crs(32750))
C.pct_sf
```

### ├ Plot the spatial object for checking

To quickly check our data, we use a few lines of `ggplot2` code to make a
*bubble map* of our variable, where the symbol *area* is proportional to the 
variable value (in this case, soil `{r} var0`, back-transformed from
log~10~).

```{r spplot-spdf, fig.height=5, fig.width=8, fig.align='center', out.width="80%", fig.cap="Visualization of spatial point data object for log10-transformed C.pct concentrations in Ridgefield A horizon soil to 2022.", message=FALSE, warning=FALSE, results='hold'}
ggplot(DTC)+
  annotation_spatial(data=rfRast) +
  geom_sf(data=rfGap, fill="#ffffff80", col="transparent") +
  geom_sf(data=DTC[DTC$DepthToClay<100,], 
          aes(size=DepthToClay), shape=21, fill="#80d0f0a0", colour="#fff")+
  geom_sf(data=DTC[DTC$DepthToClay>=100,], 
          shape=21, size=7, colour="#e0e0e0a0", fill="#ffffff40")+
  scale_size_continuous(name="Depth to Clay (cm)", breaks=c(0,20,40,60,80,100,300),
                        range=c(2,6)) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  theme_bw() + theme(legend.key = element_rect(fill="#98a088"))
```

### Calculate soil texture class for each sample

```{r get-texture-class}
rf2022tx <- data.frame(CLAY=rf2022utm$Clay.pct,SILT=rf2022utm$Silt.pct,SAND=rf2022utm$Sand.pct)
lowCSS <- which(rowSums(rf2022tx)<90)
rf2022txcat <- TT.points.in.classes(rf2022tx, text.tol=99.85, class.sys = "AU.TT")
textlist <- apply(rf2022txcat,1, function(x){which(x==max(x))})
ambiguous <- c(0)
for(i in seq_along(textlist)){
  if(length(textlist[[i]])>1) {
    ambiguous <- c(ambiguous,i)
    textlist[[i]] <- textlist[[i]][1]
  }
}
(ambiguous <- ambiguous[-1])
textnames <- names(unlist(textlist)) ; textLong <- textnames
textLong <- gsub("Cl","clay",textLong)
textLong <- gsub("Siclay","silty clay",textLong)
textLong <- gsub("SaclayLo", "sandy clay loam", textLong)
textLong <- gsub("SiclayLo","silty clay loam",textLong)
textLong <- gsub("clayLo","clay loam",textLong)
textLong <- gsub("Saclay", "sandy clay", textLong)
textLong <- gsub("SiLo","silt loam",textLong)
textLong <- gsub("SaLo","sandy loam",textLong)
textLong <- gsub("LoSa","loamy sand",textLong)
textLong <- gsub("Lo","loam",textLong)
textLong <- gsub("Sa","sand",textLong)
rf2022utm$tclass <- textLong
# length(textnames) #; head(rf2022txcat,1) ; length(textLong)
```

### Show texture classes on maps

```{r map-text-classes, fig.height=14, fig.width=10, message=FALSE, warning=FALSE}
textureA <- ggplot() +
  geom_sf(data=extent_UTM)+
  annotation_spatial(data=rfRast2) +
  geom_sf(data=rfGap, color="#00000000", fill="#00000024") +
  geom_sf(aes(shape=tclass, fill=tclass), data=rf2022utm[rf2022utm$Horiz_simp=="A",],size=3.5) +
  scale_shape_manual(values=rep(21:25,l=9), name="Soil texture") +
  scale_fill_viridis_d(option="rocket", alpha=0.7, name="Soil texture") +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0))
textureB <- ggplot() +
  geom_sf(data=extent_UTM)+
  annotation_spatial(data=rfRast2) +
  geom_sf(data=rfGap, color="#00000000", fill="#00000024") +
  geom_sf(aes(shape=tclass, fill=tclass), 
          data=rf2022utm[substr(rf2022utm$Horizon,1,1)=="B"|substr(rf2022utm$Horizon,2,2)=="B",],
          size=3.5) +
  scale_shape_manual(values=rep(21:25,l=9), name="Soil texture") +
  scale_fill_viridis_d(option="rocket", alpha=0.7, name="Soil texture") +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0))
ggpubr::ggarrange(textureA, textureB, ncol=1, nrow=2, common.legend = F,
                  legend="right")
```

### ├ Plot a map with range-class symbols

```{r TOC-tukey-plot, echo=2:19, fig.width=8.1, fig.height=7.0, fig.align='center', fig.cap="Map of C.pct concentrations expressed as symbols for concentration ranges (UQR is 75th percentile, upWx is upper whisker, med is median, LQR is 25th percentile, loWx is lower whisker). Data are from Ridgefield A horizon soils to 2022.", out.width="80%", message=FALSE, warning=FALSE,echo=FALSE}
palette(c("black",turbo(6),"grey60","grey92", "white"))
par(mar=c(4,4,1.5,1.5), mgp=c(1.4,0.3,0), lend=2, ljoin=1, tcl=-0.3, font.lab=2)
plot(st_coordinates(extent_UTM), type="n", asp=1, xaxs="i", yaxs="i",
                    xlab="Easting (UTM Zone 50, m)", 
                    ylab="Northing (UTM Zone 50, m)")
plot_tiles(rfRast, add=TRUE)
addnortharrow(pos="topleft", border=1, lwd=1, text.col=1,  
              padin=c(0.2,0.2), scale=0.8)
addscalebar(plotepsg = 32750, htin = 0.16, label.cex = 1.2, pos = "bottomleft")
# construct and plot range classes in Tukey boxplot thresholds
DTC$tukey <- cut(DTC$DepthToClay, 
    breaks = c(0,boxplot.stats(DTC$DepthToClay)$stats,999999),
    labels = c("Upper outliers","loWx-LQR","LQR-med","med-UQR","UQR-upWx","Lower outliers"))
plot(rfGap, add=TRUE, col="#00000032", border="transparent") 
# polygon(rf_boundary[,1:2], lty=3, lwd=2, border="sienna", col="#80600032")
points(DTC$geometry, pch = 21, col = 1,
       cex=(c(.5,.65,.85,1,1.4,1.8)*2)[DTC$tukey], 
       bg=seq(2,7)[DTC$tukey])
     
legend(494450, 6406850, legend=c("Upper outliers","loWsk to LQR","LQR to med",
                                 "med to UQR","UQR to upWsk","Lower outliers"), 
       cex = 1., pch=21, pt.cex=rev(c(.5,.65,.85,1,1.4,1.8))*2, inset = 0.07, 
       bg="#ffffffb0", y.intersp=1.3, pt.bg=seq(7,2,-1), x.intersp =0.9,
       text.col = 1, col = 1, box.col="#ffffff00", 
       title="DepthToClay ranges", title.font=2)
box()
```

**Table 1**: Bins for boxplot-categorised-point map in @fig:TOC-tukey-plot

```{r tukey-map-bins, echo=FALSE, results='asis'}
tukbins <- data.frame(bins=c("-outliers","loWsk to LQR","LQR to med",
                                "med to UQR","UQR to upWsk","+outliers"),
      logfrom=signif(c(-4,boxplot.stats(C.pct_sf$C.pct)$stats),3),
      logto=signif(c(boxplot.stats(C.pct_sf$C.pct)$stats,Inf),3),
      from=round(10^c(-4,boxplot.stats(C.pct_sf$C.pct)$stats),3),
      to=round(10^c(boxplot.stats(C.pct_sf$C.pct)$stats,Inf),3))
tukbins$logto <- gsub("Inf","max",tukbins$logto);tukbins$to <- gsub("Inf","max",tukbins$to)
flextuk <- flextable(tukbins) |> padding(padding=0.05, part="all") |> autofit() |> 
  flextable::align(j=2:ncol(tukbins), align="right", part="all")
htmltools_value(flextuk, ft.align = "left")
```

## Variograms and Kriging

### ├ Make a binned simple variogram object

```{r make-variog-obj, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Plot of experimental binned variogram (Cressie's semivariance with 15 bins) for C.pct in Ridgefield A horizon soils to 2022.", fig.height=3.5, fig.width=4.5, out.width="50%"}
data0 <- na.omit(DTC[,c("Easting","Northing","DepthToClay")])
variog_DTC <- variogram(DepthToClay~1, data0, cressie = TRUE) # 

par(mar = c(3,3,1,1), mgp = c(1.7,0.3,0), font.lab=2, tcl=0.3)
plot(variog_DTC$dist, variog_DTC$gamma, xlab = "Distance (m)",
     pch=19, col="blue2", 
     ylab = expression(bold(paste("Semivariance (",gamma,")"))),
     ylim = c(0,max(variog_DTC$gamma)))
```

**Table 2**: Data for @fig:make-variog-obj

```{r print-variog-obj, echo=FALSE, results='asis'}
vgr <- variog_DTC
colnames(vgr)[1:3] <- c("Number of\npoints in bin","Distance (metres)","Semivariance")
vgr[,2] <- round(vgr[,2],0)
vgr[,3] <- round(vgr[,3],3)
ftvgr <- flextable(vgr[,1:3]) |> padding(padding=0.05, part="all") |> autofit()
htmltools_value(ftvgr, ft.align = "left")
```

### ├ Fit a variogram model using weighted least squares

```{r fit variogram model, fig.height=3, fig.width=4, message=FALSE, warning=FALSE, paged.print=FALSE, results='hold'}
vmodel_DTC <- fit.variogram(variog_DTC, 
                              model = vgm(psill = NA, model = "Exp", 
                                          nugget = 0.2, 
                                          cutoff = max(variog_DTC$dist)))
vmodel_DTC
```

```{r plot-exp-model-variograms, fig.height=3.6, fig.width=5, fig.align='center', fig.cap="Plot of experimental binned variogram, and exponential variogram model, for Gravel.pct in Ridgefield A horizon soils to 2022.", out.width="50%", results='hide', echo=FALSE}
par(mar = c(4,4,1,1), mgp = c(1.7,0.3,0), tcl=0.25, font.lab=2)
with(variog_DTC, plot(dist, gamma, ylim = c(0, max(gamma)), cex.lab=1.2, 
                   lwd=2, xlab = "Distance (m)", 
                   ylab = expression(bold(paste("Semivariance (",gamma,")")))))
# plot model
x0 <- variog_DTC$dist
y0 <- vmodel_DTC$psill[1] + 
  vmodel_DTC$psill[2]*(1 - exp(-1*x0/vmodel_DTC$range[2]))
lines(x0,y0,col="#abf")
legend("bottomright", bty = "n", cex=1, pch=c(1,NA), pt.lwd=2, lwd=c(NA,1), 
       col=c(1,2), pt.cex=1.1,legend = c("Empirical variogram",
                  paste(vmodel_DTC$model[2],"variogram model")))
text(max(x0),max(y0)/2,pos=2,offset=0, cex=0.75, family='mono', col = 2,
     labels=paste(
       " nugget =",signif(vmodel_DTC$psill[1],3),"\n",
       "partial sill =", signif(vmodel_DTC$psill[2],3),"\n",
       "range =", signif(vmodel_DTC$range[2],3),"m\n",
       "practical range =",signif(vmodel_DTC$range[2]*3,3),"m" 
  ))
# NOTE factor (range -> practical range) depends on model
```

### ├ Perform kriging interpolation

### ├ first make a grid mask

```{r make grid masked to AFR sampling zones, message=FALSE, warning=FALSE, results='hold'}
(grid0 <- st_make_grid(extent_UTM, n=c(200,200), what="centers") |> st_sfc(crs=32750))

# object rfb created earlier is sf-polygon of Ridgefield boundary
rfgrid <- st_intersection(rfb, grid0)
cat("Prediction grid:\n"); rfgrid
```

### ├ Krige to grid
```{r krige-to-grid, results='hold'}
# kriged_DTC <- krige(formula = DepthToClay~1, locations = DTC, 
#                       newdata = rfgrid, model = vmodel_DTC)
# summary(kriged_DTC)
cat("_____________________________________________________________________\n\n")
idw_DTC <- idw(formula = DepthToClay ~ 1, locations = DTC, 
                      newdata = rfgrid, idp = 2)
idw_DTC$var1.var <- NULL
summary(idw_DTC)
```

### ├ Plot a map with overlay of inverse distance interpolation

```{r idw-plot, fig.width=8, fig.height=7}
par(mfrow=c(1,1), mar=c(0,4,0,1.5), oma=c(4,0,1.5,0), mgp=c(1.4,0.3,0), 
    lend=2, ljoin=1, tcl=0.3, font.lab=2)
plot(st_coordinates(extent_UTM), type="n", asp=1, xaxs="i", yaxs="i",
                    xlab="Easting (UTM Zone 50, m)", 
                    ylab="Northing (UTM Zone 50, m)")
plot_tiles(rfRast, add=TRUE)
quilt.plot(st_coordinates(st_as_sf(rfgrid))[,1], 
           st_coordinates(st_as_sf(rfgrid))[,2],idw_DTC$var1.pred,
           col = rocket(64, alpha=0.75, beg=0.2, dir=-1), add = T, horizontal=T,
          legend.lab = expression(bold("Inverse distance interpolation of Depth to Clay (cm)")), 
           legend.mar = 6, legend.cex = 1, legend.line = 1.2, text.col=10,
           axis.args = list(mgp=c(1.2,0.1,0)))
addnortharrow(pos="topleft", padin=c(0.3,0.4), scale = 0.9)
addscalebar(plotepsg = 32750, pos = "topleft")
points(rf22a[,c("Easting","Northing")], pch=16, cex=0.1, col="orchid")
box() ; mtext("(b)",3,-1.5, cex=1.5, adj=0.98)
mtext("Easting (UTM Zone 50, m)", side=1, line=2, font=2, cex=1.2)
```

### ├ Plot a map with overlay of the kriging predictions & kriging variance

```{r map-krig-mod-pred, fig.height=12.7, fig.width=7.9, fig.align='center', fig.cap="Map showing (a) predictions and (b) variance for kriging model of Gravel.pct in Ridgefield A horizon soils to 2022. Sample points are tiny dots.", message=FALSE, warning=FALSE}
par(mfrow=c(2,1), mar=c(0,4,0,1.5), oma=c(4,0,1.5,0), mgp=c(1.4,0.3,0), 
    lend=2, ljoin=1, tcl=0.3, font.lab=2)
plot(st_coordinates(extent_UTM), type="n", asp=1, xaxs="i", yaxs="i",
                    xlab="", ylab="Northing (UTM Zone 50, m)")
plot_tiles(rfRast2, add=TRUE)
axtx <- c(1,2,3,5,10,20,30,50,100,200,300,500,1000,2000,5000,10000,20000,50000)*10
quilt.plot(st_coordinates(st_as_sf(rfgrid))[,1], 
           st_coordinates(st_as_sf(rfgrid))[,2],kriged_DTC$var1.pred,
           col = rocket(64, alpha=0.7, beg=0.2, dir=-1), add = T, horizontal=T,
           legend.lab = expression(bold("Kriging prediction of Gravel.pct (mg/kg)")), 
           legend.mar = 6, legend.cex = 1, legend.line = 1.2, text.col=10,
           axis.args = list(at=log10(axtx), labels=axtx, mgp=c(1.2,0.1,0)))
addnortharrow(pos="topleft", padin=c(0.3,0.4), scale = 0.9)
addscalebar(plotepsg = 32750, pos = "topleft")
points(rf22a[,c("Easting","Northing")], pch=16, cex=0.1, col = "gold")
box() ; mtext("(a)",3,-1.5, cex=1.5,adj=0.98)

plot(st_coordinates(extent_UTM), type="n", asp=1, xaxs="i", yaxs="i",
                    xlab="Easting (UTM Zone 50, m)", 
                    ylab="Northing (UTM Zone 50, m)")
plot_tiles(rfRast2, add=TRUE)
quilt.plot(st_coordinates(st_as_sf(rfgrid))[,1], 
           st_coordinates(st_as_sf(rfgrid))[,2],kriged_DTC$var1.var,
           col = rocket(64, alpha=0.6, beg=0.2, dir=-1), add = T, horizontal=T,
           legend.lab = expression(bold("Kriging variance of Gravel.pct (mg/kg)")), 
           legend.mar = 6, legend.cex = 1, legend.line = 1.2, text.col=10,
           axis.args = list(at=log10(axtx), labels=axtx, mgp=c(1.2,0.1,0)))
addnortharrow(pos="topleft", padin=c(0.3,0.4), scale = 0.9)
addscalebar(plotepsg = 32750, pos = "topleft")
points(rf22a[,c("Easting","Northing")], pch=16, cex=0.1, col="orchid")
box() ; mtext("(b)",3,-1.5, cex=1.5, adj=0.98)
mtext("Easting (UTM Zone 50, m)", side=1, line=2, font=2, cex=1.2)
```

--------------------------------------------------------------------------------

## Depth profiles

```{r map-landforms, fig.height=7.4, fig.width=10, fig.align='center', fig.cap="Map of Ridgefield Farm and adjacent area showing simplified landforms, with '+' symbols showing sample locations. Raster generated using the maptiles R package, with ESRI World Imagery tiles.", out.width="75%", warning=FALSE, message=FALSE, results='hide'}
colors <- c("Residual"="#80000080", Erosional="#F0E68C80", Depositional="#36648B80")
ggplot()+
  geom_sf(data=extent_UTM) +
  annotation_spatial(data=rfRast) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x="Easting (m)", y="Northing (m)", 
       caption="CRS: UTM Zone 50 South (EPSG:32750)", color="Legend") +
  # geom_sf(data=rfGap, fill="#ffffff40", col="transparent") +
  geom_sf(aes(fill="Residual"), data=RF_Res_sf, color="grey") +
  geom_sf(aes(fill="Erosional"), data=RF_Ero_sf, color="grey") +
  geom_sf(aes(fill="Depositional"), data=RF_Dep_sf, color="grey") +
  geom_sf(data=rf22A.utm, shape=3, col="#fff") +
  annotation_north_arrow(location="bl", pad_x=unit(0.02,"cm"), pad_y=unit(.02,"cm"),
                         style=north_arrow_fancy_orienteering(text_size=14)) +
  annotation_scale(location="bl", text_cex=1.1, pad_x=unit(2,"cm")) +
  theme_bw() + theme(axis.title=element_text(size=14, face="bold"),
                     legend.key=element_rect(fill="#8c8c85"),
                     legend.title = element_text(color="transparent"),
                     legend.text=element_text(size=12)) +
  scale_fill_manual(values=colors) +
  coord_sf(datum=st_crs(32750))
```

### create depth and landform categories

```{r}
rf2022$Depth.categ <- cut(rf2022$Depth.mean, breaks=seq(0,100,10), 
                          labels=as.character(seq(5,95,10)))
rf2022$Depth.cat.num <- as.numeric(as.character(rf2022$Depth.categ))
rf22utm <- st_as_sf(rf2022[!is.na(rf2022$Northing),], 
                    coords=c("Easting","Northing"), crs=32750, remove=F)
rf22utm$LF3 <- rep(NA,nrow(rf22utm))
rf22utm$LF3[unlist(st_contains(RF_Res_sf,rf22utm))] <- "Residual"
rf22utm$LF3[unlist(st_contains(RF_Ero_sf,rf22utm))] <- "Erosional"
rf22utm$LF3[unlist(st_contains(RF_Dep_sf,rf22utm))] <- "Depositional"
rf22utm$LF3 <- factor(rf22utm$LF3, levels=c("Residual","Erosional","Depositional"))
rf22res <- rf22utm[rf22utm$LF3=="Residual",]
rf22res$LF3 <- droplevels(rf22res$LF3)
rf22ero <- rf22utm[rf22utm$LF3=="Erosional",] 
rf22ero$LF3 <- droplevels(rf22ero$LF3)
rf22dep <- rf22utm[rf22utm$LF3=="Depositional",]
rf22dep$LF3 <- droplevels(rf22dep$LF3)
```

```{r depth-profiles-SOC-CNratio, fig.height=4.5, fig.width=6, fig.cap="Depth profiles for soil organic carbon (SOC) and carbon to nitrogen ratio, separated by simplified landform."}
C.pct.res <- with(rf22res, tapply(C.pct, Depth.categ, mean, na.rm=T))
C.pct.ero <- with(rf22ero, tapply(C.pct, Depth.categ, mean, na.rm=T))
C.pct.dep <- with(rf22dep, tapply(C.pct, Depth.categ, mean, na.rm=T))

par(mfrow=c(1,2), mar=c(1,3,3,1), mgp=c(1.7,0.3,0), tcl=0.25, font.lab=2, las=1)
palette(c("black",viridis(10),"white","transparent"))
plot(seq(5,95,10) ~ C.pct.res, type="b", xlim=c(0,5), ylim=c(100,0), xaxt="n",  
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", 
     xaxs="i", yaxs="i")
# lines(seq(5,95,10) ~ C.pct.res, type="c")
axis(3);mtext("SOC (%)", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ C.pct.ero, type="b", ylim=c(100,0), xaxt="n",  
     pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ C.pct.dep, type="b", ylim=c(100,0), xaxt="n", 
     pch=24, col="#248", bg="#36648B80")
legend("bottomright", bty="n", legend=c("Residual","Erosional","Depositional"),
       pt.bg=13, col=c("#500","#874","#248"),lwd=1)
legend("bottomright", bty="n", legend=c("Residual","Erosional","Depositional"),
       pch=c(21,22,24), col=c("#600","#a95","#469"),
       pt.bg=c("#BF7F7F","#F7F2C5","#9AB1C5"), x.intersp=2)

C.N.ratio.res <- with(rf22res, tapply(C.N.ratio, Depth.categ, mean, na.rm=T))
C.N.ratio.ero <- with(rf22ero, tapply(C.N.ratio, Depth.categ, mean, na.rm=T))
C.N.ratio.dep <- with(rf22dep, tapply(C.N.ratio, Depth.categ, mean, na.rm=T))

plot(seq(5,95,10) ~ C.N.ratio.res, type="b", xlim=c(5,29), ylim=c(100,0),   
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", 
     xaxt="n", xaxs="i", yaxs="i")
axis(3);mtext("C:N Ratio", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ C.N.ratio.ero, type="b", ylim=c(100,0), xaxt="n",  
     pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ C.N.ratio.dep, type="b", ylim=c(100,0), xaxt="n", 
     pch=24, col="#248", bg="#36648B80")
```

```{r calcs-for-depth-profile-xlim, eval=FALSE, message=FALSE, warning=FALSE}
cat("min:",min(cbind(Gravel.pct.res,
          Gravel.pct.ero,
          Gravel.pct.dep),
    na.rm=T),"max:",
max(cbind(Gravel.pct.res,
          Gravel.pct.ero,
          Gravel.pct.dep),
    na.rm=T))
```

```{r depth-profiles-clay-gravel, fig.height=4.5, fig.width=6, echo=FALSE, fig.cap="Depth profiles for clay and gracvel content, separated by simplified landform."}
Clay.pct.res <- with(rf22res, tapply(Clay.pct, Depth.categ, mean, na.rm=T))
Clay.pct.ero <- with(rf22ero, tapply(Clay.pct, Depth.categ, mean, na.rm=T))
Clay.pct.dep <- with(rf22dep, tapply(Clay.pct, Depth.categ, mean, na.rm=T))

par(mfrow=c(1,2), mar=c(1,3,3,1), mgp=c(1.7,0.3,0), tcl=0.25, font.lab=2, las=1)
palette(c("black",viridis(10),"white","transparent"))
plot(seq(5,95,10) ~ Clay.pct.res, type="b", xlim=c(5,35), ylim=c(100,0), xaxt="n",  
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", yaxs="i")
# lines(seq(5,95,10) ~ Clay.pct.res, type="c")
axis(3);mtext("Clay (%)", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ Clay.pct.ero, type="b", pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ Clay.pct.dep, type="b", pch=24, col="#248", bg="#36648B80")

Gravel.pct.res <- with(rf22res, tapply(Gravel.pct, Depth.categ, mean, na.rm=T))
Gravel.pct.ero <- with(rf22ero, tapply(Gravel.pct, Depth.categ, mean, na.rm=T))
Gravel.pct.dep <- with(rf22dep, tapply(Gravel.pct, Depth.categ, mean, na.rm=T))

plot(seq(5,95,10) ~ Gravel.pct.res, type="b", xlim=c(11,30), ylim=c(100,0),   
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", 
     xaxt="n", yaxs="i")
axis(3);mtext("Gravel (%)", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ Gravel.pct.ero, type="b", pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ Gravel.pct.dep, type="b", pch=24, col="#248", bg="#36648B80")
legend("bottomleft", bty="n", legend=c("Residual","Erosional","Depositional"),
       pch=c(21,22,24), col=c("#600","#a95","#469"),
       pt.bg=c("#BF7F7F","#F7F2C5","#9AB1C5"), cex=0.9)
```

```{r depth-profiles-pH-EC, fig.height=4.5, fig.width=6, echo=FALSE, fig.cap="Depth profiles for soil pH (CaCl₂) and EC, separated by simplified landform."}
pH.CaCl2.res <- with(rf22res, tapply(pH.CaCl2, Depth.categ, mean, na.rm=T))
pH.CaCl2.ero <- with(rf22ero, tapply(pH.CaCl2, Depth.categ, mean, na.rm=T))
pH.CaCl2.dep <- with(rf22dep, tapply(pH.CaCl2, Depth.categ, mean, na.rm=T))

par(mfrow=c(1,2), mar=c(1,3,3,1), mgp=c(1.7,0.3,0), tcl=0.25, font.lab=2, las=1)
palette(c("black",viridis(10),"white","transparent"))
plot(seq(5,95,10) ~ pH.CaCl2.res, type="b", xlim=c(4,6), ylim=c(100,0), xaxt="n",  
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", 
     xaxs="i", yaxs="i")
# lines(seq(5,95,10) ~ pH.CaCl2.res, type="c")
axis(3);mtext("SOC (%)", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ pH.CaCl2.ero, type="b", ylim=c(100,0), xaxt="n",  
     pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ pH.CaCl2.dep, type="b", ylim=c(100,0), xaxt="n", 
     pch=24, col="#248", bg="#36648B80")

EC.res <- with(rf22res, tapply(EC, Depth.categ, mean, na.rm=T))
EC.ero <- with(rf22ero, tapply(EC, Depth.categ, mean, na.rm=T))
EC.dep <- with(rf22dep, tapply(EC, Depth.categ, mean, na.rm=T))

plot(seq(5,95,10) ~ EC.res, type="b", ylim=c(100,0), xlim=c(0,2200),   
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", 
     xaxt="n", xaxs="i", yaxs="i")
axis(3);mtext("C:N Ratio", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ EC.ero, type="b", ylim=c(100,0), xaxt="n",  
     pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ EC.dep, type="b", ylim=c(100,0), xaxt="n", 
     pch=24, col="#248", bg="#36648B80")
legend("topright", bty="n", legend=c("Residual","Erosional","Depositional"),
       pt.bg=13, col=c("#500","#874","#248"),lwd=1, cex=0.9)
legend("topright", bty="n", legend=c("Residual","Erosional","Depositional"),
       pch=c(21,22,24), col=c("#600","#a95","#469"),
       pt.bg=c("#BF7F7F","#F7F2C5","#9AB1C5"), x.intersp=2, cex=0.9)
```

```{r depth-profiles-exAl-exNa, fig.height=4.5, fig.width=6, echo=FALSE, fig.cap="Depth profiles for exchangeable Al and exchangeable Na, separated by simplified landform."}
ex.Al.res <- with(rf22res, tapply(ex.Al, Depth.categ, mean, na.rm=T))
ex.Al.ero <- with(rf22ero, tapply(ex.Al, Depth.categ, mean, na.rm=T))
ex.Al.dep <- with(rf22dep, tapply(ex.Al, Depth.categ, mean, na.rm=T))

par(mfrow=c(1,2), mar=c(1,3,3,1), mgp=c(1.7,0.3,0), tcl=0.25, font.lab=2, las=1)
palette(c("black",viridis(10),"white","transparent"))
plot(seq(5,95,10) ~ ex.Al.res, type="b", ylim=c(100,0), xaxt="n", xlim=c(0,.6), 
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", 
     xaxs="i", yaxs="i")
# lines(seq(5,95,10) ~ ex.Al.res, type="c")
axis(3);mtext("Exchangeable Al", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ ex.Al.ero, type="b", ylim=c(100,0), xaxt="n",  
     pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ ex.Al.dep, type="b", ylim=c(100,0), xaxt="n", 
     pch=24, col="#248", bg="#36648B80")

ex.Na.res <- with(rf22res, tapply(ex.Na, Depth.categ, mean, na.rm=T))
ex.Na.ero <- with(rf22ero, tapply(ex.Na, Depth.categ, mean, na.rm=T))
ex.Na.dep <- with(rf22dep, tapply(ex.Na, Depth.categ, mean, na.rm=T))

plot(seq(5,95,10) ~ ex.Na.res, type="b", ylim=c(100,0), xlim=c(0,4.5),  
     pch=21, col="#500", bg="#80000080", ylab = "Mean depth (cm)", 
     xaxt="n", xaxs="i", yaxs="i")
axis(3);mtext("Exchangeable Na", side=3, line=1.7, font=2)
points(seq(5,95,10) ~ ex.Na.ero, type="b", ylim=c(100,0), xaxt="n",  
     pch=22, col="#874", bg="#F0E68C80")
points(seq(5,95,10) ~ ex.Na.dep, type="b", ylim=c(100,0), xaxt="n", 
     pch=24, col="#248", bg="#36648B80")
legend("bottomright", bty="n", legend=c("Residual","Erosional","Depositional"),
       pt.bg=13, col=c("#500","#874","#248"),lwd=1, inset=c(0,0.05))
legend("bottomright", bty="n", legend=c("Residual","Erosional","Depositional"),
       pch=c(21,22,24), col=c("#600","#a95","#469"),
       pt.bg=c("#BF7F7F","#F7F2C5","#9AB1C5"), x.intersp=2, inset=c(0,0.05))
```

